Smelt Cage Histopathology

### 2023-10-19 recaptureanalysis/recaps.Rmd
reorganized repo

```{r setup, include=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE)
library(tidyverse)
library(readxl)
library(sf)
library(lubridate)
library(spacetools)
library(deltamapr)
#load("Recaps.RData")
```

@@

recaps = read_excel("Recaptured Experimental Release_2022-2023_working draft.xlsx", 
                    sheet = "Adult Recaptures") %>%
  rename(Site = `Release Site`, Type = `Release Type`, Latitude= `Start Latitude`, Longitude = `Start Longitude`)

recaps    =     mutate(recaps, fishID = paste("smelt", 1:nrow(recaps)),
                Type = case_when(Type == "header" ~ "Soft",
                                 TRUE ~ Type)) %>%
  rename(Adclip = `Ad clip`, Mark = `VIE Color/Position`) %>%
  select(-Type, -Site)
str(recaps)

@@

recaps3
```

@@

```{r}
ggplot(recaps3, aes(x = SiteType, y = Distance))+ geom_boxplot(aes(fill = Type))+
  ylab("Distance from Release Site (m)")+
  theme_bw()
```

@@

recapsum

@@

I'm not quite sure the best way to summarize this. I'll start by just the mean of the distance matrix, but i might want do just half the matrix. I'll look into that later.

@@

nndist = bind_rows(RVdtx, RVdt, SCs, SCh)

ggplot(nndist, aes(x = Release, y = Distance, fill = Release)) + geom_boxplot()+
  ylab("Distance Between Recaptures")

@@

Now let's bring in the time aspect

@@

recapstime = read_excel("RecapturesSummary_1.xlsx", sheet = "REcapture Distances")

@@

>>>>

```{r setup, include=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE)
library(tidyverse)
library(readxl)
library(sf)
library(lubridate)
library(spacetools)
library(deltamapr)
load("Recaps.RData")
#save(nndist, distances, recaps, recaps2, recaps3, RVhdist, SCsoft, SChard, RVhdistx, Dwschard, Dwscsoft, file = "Recaps.RData")
```

@@

recaps = read_excel("Recaptured Experimental Release_2022-2023_working draft (2).xlsx", 
                    sheet = "Full DSM Data") %>%
  rename(Site = `Release Site`, Type = `Release Type`, Latitude= `Start Latitude`, Longitude = `Start Longitude`,
         Mark = `VIE Color/Position`)
recaps    =     mutate(recaps, fishID = paste("smelt", 1:nrow(recaps)),
                Type = case_when(Type == "header" ~ "Soft",
                                 is.na(Type) ~ "Wild",
                                 TRUE ~ Type),
                Site = case_when(is.na(Site) ~ "Wild",
                                 TRUE ~ Site))
str(recaps)

@@

recaps3

write.csv(recaps, "Recapturedistances.csv")
```

@@

```{r}
ggplot(filter(recaps3, Type != "Hard (Trailer)", Type != "Wild"), aes(x = SiteType, y = Distance))+ 
  geom_boxplot(aes(fill = SiteType))+
  ylab("Distance from Release Site (m)")+
  scale_fill_discrete(guide = NULL)+
  theme_bw()
```

@@

recapsum
```

#ANOVA of differences in distance by release group
library(emmeans)
m1 = lm(Distance ~ SiteType, data = recaps3)
summary(m1)
anova(m1)
emmeans(m1, pairwise ~ SiteType)
plot(emmeans(m1, pairwise ~ SiteType), comparisons = T)

#so hard releases at RIo Vista traveled further from hard releases at SDWSC, but no other significant differences

```
```{r}

#ANOVA of differences in distance by release event
m1.1 = lm(Distance ~ ReleaseEvent, data = recaps3)
summary(m1.1)
anova(m1.1)
emmeans(m1.1, pairwise ~ ReleaseEvent)
plot(emmeans(m1.1, pairwise ~ ReleaseEvent), comparisons = T)

#It's a bit of a mess.

Plot of distance to recapture over time by release type

ggplot(recaps3, aes(x = as.numeric(DaysSinceRelease), y = Distance, color = Type))+
  geom_point()+ geom_smooth(method = "lm")+ ylab("Distance from Release Site (m)")+
  xlab("Days since release")

Plot of distance to recapture over time by release site

ggplot(filter(recaps3, Type == "Hard"), aes(x = as.numeric(DaysSinceRelease), y = Distance, color = Site))+
  geom_point()+ geom_smooth(method = "lm")+ ylab("Distance from Release Site (m)")+
  xlab("Days since release")

@@

Let's look at the mean of the lower triagle of each of these.

@@

nndist = bind_rows(RVdtx, RVdt, SCs, SCh) %>%
  mutate(Release = factor(Release, levels = c("Rio Vista Hard", "DWSC Hard", "DWSC Soft", "Trailer"),
                          labels = c("Hard Rio Vista", "Hard SDWSC", "Soft SWDSC", "Trailer")))

ggplot(filter(nndist, Release != "Trailer"), aes(x = Release, y = Distance, fill = Release)) + geom_boxplot()+
  ylab("Distance Between Recaptures")+ theme_bw()+ scale_fill_discrete(guide = NULL)

@@

now let's look at it statistically

@@

m2 = lm(Distance ~ Release, data = nndist)
summary(m2)
plot(m2)
#QQ plot is a bit of a mess I tried log-transforming and sqrt transforming and its still a mess.

emmeans(m2, pairwise ~ Release)
plot(emmeans(m2, pairwise ~ Release), comparisons = T)

@@

So DWSC soft were closer together than rio vista hard, but no difference between DWSC hard and DWSC soft

I could also do this by release rather than release type

```{r}
smeltIDS = select(recaps2, fishID, ReleaseEvent)

nndist2 = left_join(nndist, smeltIDS, by = c("Fish1" = "fishID"))

m3 = lm(Distance ~ ReleaseEvent, data = nndist2)
summary(m3)
plot(m3)
#QQ plot is a bit of a mess I tried log-transforming and sqrt transforming and its still a mess.

emmeans(m3, pairwise ~ ReleaseEvent)
plot(emmeans(m3, pairwise ~ ReleaseEvent), comparisons = T)

write.csv(nndist2, "DistanceMatrix.csv")
```

### 2023-10-19 smelt_2019_allseasons/histopathology.Rmd
reorganized repo

```{r setup, include=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(lubridate)
library(readxl)
library(car)
library(emmeans)
library(effects)
library(ordinal)
library(sure)

>>>>

```{r setup, include=FALSE, warning=FALSE}

#making a change to test branching

knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(lubridate)
library(readxl)
library(car)
library(emmeans)
library(effects)
library(ordinal)
library(sure)
